---
title: "Bivariate Visualizations"
author: "Brianna Heggeseth"
logo: "images/mac.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: true
    transition: fade
    chalkboard: true
    auto-stretch: false
    scrollable: true
editor: visual
---

```{r}
#| include: false
#| label: 4_setup

knitr::opts_chunk$set(echo = FALSE, error = TRUE)
library(tufte)
library(tidyverse)
library(ggplot2)
library(fivethirtyeight)
library(readr)
```

## Learning Goals {.unnumbered}

-   Identify appropriate types of bivariate visualizations, depending on the type of variables (categorical, quantitative)
-   Create basic bivariate visualizations based on real data


## Bivariate Visualizations {.unnumbered}

In this activity we will analyze data from the 2016 presidential election.

We'll explore county-level election outcomes and demographics.

## Template File {.unnumbered}

Download a template .Rmd of this activity. Put the file in a `Day_04` folder within your `COMP_STAT_112` folder.

-   This .Rmd only contains exercises that we'll work on in class and you'll finish for Assignment 3.

## Getting to know the dataset {.smaller}

Begin by loading the data from https://www.macalester.edu/\~ajohns24/data/electionDemographics16.csv and getting to know the data.

## Getting to know the dataset {.unnumbered .smaller}

Loading in the Data

```{r echo=TRUE}
elect <- read_csv("https://www.macalester.edu/~ajohns24/data/electionDemographics16.csv")
```

## Getting to know the dataset {.smaller}

Check out the first rows of elect. What are the units of observation?

. . .

```{r echo=TRUE}
head(elect)
```

## Getting to know the dataset {.smaller}

How much data do we have?

. . .

```{r echo=TRUE}
dim(elect)
```

## Getting to know the dataset {.smaller}

What are the names of the variables?

. . .

```{r echo=TRUE}
names(elect)
```

## Review: Univariate Viz {.smaller}

Categorical Variable: Bar Plot

```{r echo=TRUE}
ggplot(data = elect, aes(x = winrep_2016)) + 
    geom_bar()
```

## Review: Univariate Viz {.smaller}

Quantitative Variable: Histogram or Density plot

```{r echo=TRUE}
ggplot(data = elect, aes(x = perrep_2016)) + 
    geom_histogram()

ggplot(data = elect, aes(x = perrep_2016)) + 
    geom_density()
```

## Preview: Bivariate Viz {.smaller}

Quantitative + Quantitative Variable: Scatterplot

```{r}
ggplot(data = elect, aes(x = perrep_2012, y = perrep_2016)) + 
    geom_text(aes(label = abb)) 
```

## Preview: Bivariate Viz {.smaller}

Quantitative + Categorical Variable: Density Plots, Boxplots, etc.

```{r}
ggplot(data = elect, aes(fill = StateColor, x = perrep_2016)) + 
    geom_density(alpha = 0.5) +
    scale_fill_manual(values = c("blue", "purple", "red"))
```

## Preview: Bivariate Viz {.smaller}

Categorical + Categorical Variable: side-by-side, proportion Bar plots, etc.

```{r}
ggplot(elect, aes(x = StateColor, fill = winrep_2016)) +
  geom_bar(position = "fill")
```

## In Class

Work on the activity, checking in with your mates at your table. 

Notice patterns! Feel free to make visualizations more effective as you go along.

## After Class

You'll make sure to complete Exercise 8-17 for the Assignment 3 (due next Tues).

For Friday's class, meet in the Library (Idea Lab for morning, Lib 206 for FYC)!