---
title: "Factors"
author: "James Normington"
logo: "images/mac.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: true
    transition: fade
    chalkboard: true
    auto-stretch: false
editor: visual
---

##  {.unnumbered .smaller}

## Announcements {.unnumbered .smaller}

-   No class Thursday -- MSCS Capstone Days
    -   Assignment 7 -- attend two talks and write a reflection. Description on course schedule.
-   Midterm next Thursday!
    -   Link to description/study guide on course schedule

    -   Next Tuesday we will review

## Template File {.unnumbered}

Download a template .Rmd of this activity. Put the file in a `Day_10` folder within your `COMP_STAT_112` folder.

-   This .Rmd only contains 1 exercise
-   Submit today's activity as Assignment 6 Part 2

## Factors vs. Characters {.unnumbered}

Both: can represent categorical variables

character: ***does not*** have predefined levels

-   Names

-   Addresses

-   Free responses (e.g., How does this class make you feel?)

## Factors vs. Characters {.unnumbered}

R class factor: ***does*** have predefined levels, better for categorical analysis:

-   Final grade: A, A-, B+, B....

-   Favorite season: winter, spring, summer, autumn

-   Multiple choice response (e.g., How do you feel about this class? `love`, `like`, `neutral`, `dislike`, `hate`)

## Factors vs. Characters in R {.unnumbered}

Variables with class `character` do not have `level`s

```{r, include = TRUE, echo = TRUE}
library(readr); library(dplyr)
Grades <- read_csv("https://jamesnormington.github.io/112_spring_2023/data/grades.csv")

head(Grades)
class(Grades$grade)
levels(Grades$grade)

```

## Converting a character to factor {.unnumbered}

```{r, include = TRUE, echo = TRUE}
Grades = Grades %>%
  mutate(grade = factor(grade))

levels(Grades$grade)
```

## Changing the order of levels {.unnumbered}

1.  `fct_relevel()`: manually reorder levels
2.  `fct_infreq()`: order levels from higher to lowest frequency
3.  `fct_reorder()`: reorder levels by values of another variable
4.  `fct_rev()`: reverse current order

## Changing the value of levels {.unnumbered}

1.  `fct_recode():` manually change levels

2.  `fct_lump():` group together least common levels

## Work time {.unnumbered}

Work on the short Factor examples, then go on to Assignment 6 or study for Midterm.

1.  Work in groups!
2.  Ask me questions!
