---
title: "Catch up & Iterative Visualizations"
author: "Brianna Heggeseth"
logo: "images/mac.png"
format: 
  revealjs:
    theme: slides.scss
    multiplex: true
    transition: fade
    chalkboard: true
    auto-stretch: false
editor: visual
---

## Announcements {.unnumbered .smaller}

**This week in MSCS**

-   Next Wednesday (4:45-7pm): MSCS Block Party! Outside OLRI near Tennis Courts

**Class Masking Policy (Starting Today)**

-   Section 1: Masks required
-   Section 2: Masks optional
-   FYC: Masks optional

**Taking Care of Each Other**

-   If you are not feeling well, stay home/in the dorm & rest.
-   If you are exposed or a bit unwell, wear a mask.
-   Masks are always welcome!


## Sticky Points {.unnumbered .smaller}

This is the last *formal* day of visualizations before moving on to data wrangling.

What aspects are you getting stuck on?

- Discuss in groups
- Offer tips/resources to each other
- Share common sticky points with class

## Monthly Learning Reflections {.unnumbered .smaller}

To practice self-assessment and self-reflection, I would like you to complete a monthly self-reflection about your learning in this class:

- by the end of Sept
- by the end of Oct
- by the end of Nov

Open the [template file](https://docs.google.com/document/d/1ZpU8y9rOH_lRvZsVWeq-WEx_qwzSkKT484l9QxphJjU/edit?usp=sharing). 

- What question do you have?

*I'll create individual versions of this document and share them with you. You can add your reflections in your individual document. This will only be shared between you and me.*

## ggplot review {.unnumbered .smaller}

For each function, tell me what it does and what types of input arguments it takes.

- `ggplot()`
- `geom_point()`
- `aes()`
- `scale_color_manual()`
- `ggmap()`
- `geom_histogram()`
- `geom_density()`
- `labs()`



## ggplot example {- .smaller .scrollable}

```{r echo=FALSE}
library(tidyverse)
library(fivethirtyeight)
data(bechdel)
newbechdel <- bechdel %>%
  mutate(clean_test = factor(bechdel$clean_test, c("nowomen", "notalk", "men", "dubious", "ok"))) %>%
  mutate(yearCat = cut(year, breaks = seq(1969, 2014, by = 5)))


```

```{r echo=TRUE, fig.width=12, fig.height=6 }
yvalues <- newbechdel %>%
  group_by(yearCat) %>%
  count(pass = clean_test %in% c('dubious','ok')) %>%
  mutate(prop = n/sum(n)) %>% filter( pass == TRUE) %>% pull(prop)

bars2 <- data.frame(bars = c(sapply(yvalues, function(x) c(NA,x,rep(NA,3)))) ,yearCat = rep(levels(newbechdel$yearCat),each=5), clean_test = rep(levels(newbechdel$clean_test), 9))

ggplot(newbechdel, 
       aes(
         x = yearCat, 
         fill = clean_test)) + 
  geom_bar(
    position = "fill",
    width = .98) + 
  scale_fill_manual(
    values = c("red","salmon","pink","steelblue1","steelblue4"),
    guide = 'none') +
  labs(
    caption = 'SOURCE: BECHDELTEST.COM', 
    title='Bechdel Test Over Time', 
    subtitle = 'How women are represented in movies') + 
  scale_y_continuous(
    name = '',
    breaks = seq(0,1,by=.25),
    labels = scales::percent) +
  scale_x_discrete(
    name = '',
    labels = c("1970-\n'74","","1980-\n'84","","1990-\n'94","","2000-\n'05","","2010-\n'13")) +
  geom_errorbar(
    data = bars2, 
    mapping = aes(y=bars,ymin=bars,ymax=bars),
    color = "black",
    size = 2) +
  annotate(
    'text',
    x = 9.6,
    y = rev(c(0.25,.5,.6,.8,.97)),
    label = c('Fewer than\n two women',"Women don't\n talk to each\n other", 'Women only\n talk about men','Dubious','Passes\n Bechdel\n Test'),
    hjust = 0) +
  annotate(
    'text',
    x = c(5,5),
    y = c(.25,.75),
    label=c('PASS','FAIL'),
    hjust = .5, 
    size=20) + 
  coord_cartesian(
    xlim=c(1,10), 
    clip="off") + 
  theme_classic() 
```




## Iterative Viz {.unnumbered .smaller}

[Iterative Viz Instructions](https://docs.google.com/document/d/1u_Ca39P75W77gxv1lq3KhYeyT_Mfy7Qg18E-lJBCFeU/edit?usp=sharing)

Steps:

1. Complete TidyTuesday 1, 2, 3, or 4!
2. Submit one of those graphics (no need to make changes) to Viz 0 assignment by Sept 30
3. Revise/update your graphic based on feedback and submit for Viz 1 by Oct 31
4. Revise/update your graphic based on feedback and submit for Viz 2 by Nov 30

Creativity is highly encouraged and increased complexity (that increases the effectiveness of the visual) is expected.


## Rest of Class {.unnumbered .smaller}

Choose your work:

- Assignments 4 (due Sunday) or 5 (Tuesday)
- Tidy Tuesday 3 (due today)
